### 2018-04-05 concepts.tex
figure for types of determinism

>>>>

\begin{figure}
\includegraphics[width=\columnwidth]{types}
\caption{Types of determinism}
\label{fig:types}
\end{figure}

We define two basic types of determinism, shown in Figure
\ref{fig:types}:  horizontal determinism and vertical determinism.  In
horizontal determinism, which is what we usually think of when we
think about deterministic behavior, a software system reliably
produces the same behavior given the same steps of a test: the
behavior of multiple executions that are ``the same'' in terms of
inputs/actions can be aligned and checked for some type of equality.  In
vertical determinism, rather than behavior across executions, we are
interested in behavior within an execution, where repeating the same
step of a test twice should result in the same behavior.  Obviously,
vertical determinism is both rarer and more specialized than
horizontal determinism.  However, in some instances, as we discuss
below, vertical determinism can be extremely important, both for
correctness and for ease-of-use of software interfaces.
